---
title: "Creating SIR Models with R6Sim"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Creating SIR Models with R6Sim}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE
)
```

## Introduction

This vignette demonstrates how to create a basic epidemiological model using the `R6Sim` framework. We'll implement the classic SIR (Susceptible-Infectious-Recovered) model using the `deSolve` package to solve the differential equations.

## The SIR Model

The SIR model divides a population into three compartments:
- **S**: Susceptible individuals who can become infected
- **I**: Infectious individuals who can transmit the disease
- **R**: Recovered (or removed) individuals who are immune or deceased

The model is governed by these differential equations:

$$\frac{dS}{dt} = -\beta \frac{SI}{N}$$
$$\frac{dI}{dt} = \beta \frac{SI}{N} - \gamma I$$
$$\frac{dR}{dt} = \gamma I$$

Where:
- $\beta$ is the transmission rate
- $\gamma$ is the recovery rate
- $N$ is the total population size

## Implementation with R6Sim

First, let's create an R6 class that inherits from R6Sim:

```{r setup}
library(R6Sim)
library(dplyr)
library(magrittr)
library(deSolve)

# Create an SIR epidemic model
SIRModel <- R6::R6Class(
  classname = "SIRModel",
  inherit = R6Sim,
  
  public = list(
    # For storing results
    results = NULL,
    
    # Initialize the model
    initialize = function(name) {
      super$initialize(name = name)
      
      # Set default parameter values
      self$set_input("N", 10000, "settings")$          # Total population
        set_input("I0", 10, "settings")$               # Initial infected
        set_input("R0", 0, "settings")$                # Initial recovered
        set_input("beta", 0.3, "natural_history")$     # Transmission rate
        set_input("gamma", 0.1, "natural_history")$    # Recovery rate
        set_input("t.max", 150, "settings")$           # Time horizon (days)
        set_input("dt", 1, "settings")                 # Time step (days)
    },
    
    # Required setup method
    setup_run = function() {
      # Any pre-simulation setup
      invisible(self)
    },
    
    # SIR model equations
    sir_equations = function(time, state, parameters) {
      with(as.list(c(state, parameters)), {
        # Compute the derivatives
        dS <- -beta * S * I / N
        dI <- beta * S * I / N - gamma * I
        dR <- gamma * I
        
        # Return the derivatives
        return(list(c(dS, dI, dR)))
      })
    },
    
    # Core simulation method
    simulate = function() {
      # Initial conditions
      initial_state <- c(
        S = self$inputs$N - self$inputs$I0 - self$inputs$R0,
        I = self$inputs$I0,
        R = self$inputs$R0
      )
      
      # Parameters for the ODE solver
      parameters <- c(
        N = self$inputs$N,
        beta = self$inputs$beta,
        gamma = self$inputs$gamma
      )
      
      # Time points
      times <- seq(0, self$inputs$t.max, by = self$inputs$dt)
      
      # Solve the ODEs
      self$results <- as.data.frame(ode(
        y = initial_state,
        times = times,
        func = self$sir_equations,
        parms = parameters
      ))
      
      # Calculate key epidemic metrics
      peak_infected <- max(self$results$I)
      peak_day <- self$results$time[which.max(self$results$I)]
      final_size <- self$results$R[length(self$results$R)]
      attack_rate <- final_size / self$inputs$N
      
      # Return summary statistics
      return(data.frame(
        peak_infected = peak_infected,
        peak_day = peak_day,
        final_size = final_size,
        attack_rate = attack_rate,
        R0 = self$inputs$beta / self$inputs$gamma
      ))
    },
    
    # Plot epidemic curves
    plot_curves = function() {
      if (is.null(self$results)) {
        stop("You need to run simulate() first")
      }
      
      # Reshape for plotting
      plot_data <- tidyr::pivot_longer(
        self$results, 
        cols = c("S", "I", "R"),
        names_to = "Compartment", 
        values_to = "Count"
      )
      
      return(plot_data)
    }
  )
)

# Create an instance of the model
sir_model <- SIRModel$new("basic_sir_model")
```

## Running the Model

Now let's run the model with the default parameters:

```{r}
# Run the model
sir_model$setup_run()
results <- sir_model$simulate()
results
```

## Visualizing the Results

We can visualize the classic SIR curves:

```{r, fig.width=7, fig.height=5}
# Get the epidemic curves data
curves_data <- sir_model$plot_curves()

# Plot using base R or ggplot2 if available
if (requireNamespace("ggplot2", quietly = TRUE)) {
  library(ggplot2)
  
  ggplot(curves_data, aes(x = time, y = Count, color = Compartment)) +
    geom_line(linewidth = 1) +
    scale_color_manual(values = c("I" = "red", "R" = "green", "S" = "blue")) +
    theme_minimal() +
    labs(title = "SIR Model Dynamics",
         x = "Time (days)",
         y = "Population")
} else {
  # Fallback to base R plotting
  plot(sir_model$results$time, sir_model$results$S, type = "l", col = "blue",
       xlab = "Time (days)", ylab = "Population", 
       main = "SIR Model Dynamics",
       ylim = c(0, sir_model$inputs$N))
  lines(sir_model$results$time, sir_model$results$I, col = "red")
  lines(sir_model$results$time, sir_model$results$R, col = "green")
  legend("right", legend = c("Susceptible", "Infected", "Recovered"),
         col = c("blue", "red", "green"), lty = 1)
}
```

## Exploring Different Parameters

Let's see how changing the transmission rate affects the epidemic:

```{r}
# Change the transmission rate and re-run
sir_model$set_input("beta", 0.5, "natural_history")
sir_model$setup_run()
results_high_beta <- sir_model$simulate()

# Compare results
results
results_high_beta
```

We can also explore changes in the recovery rate:

```{r}
# Change the recovery rate and re-run
sir_model$set_input("gamma", 0.2, "natural_history")
sir_model$setup_run()
results_high_gamma <- sir_model$simulate()

# Compare all results
rbind(
  "Original" = results[1,],
  "High Beta" = results_high_beta[1,],
  "High Gamma" = results_high_gamma[1,]
)
```

## Parameter Distributions for Uncertainty Analysis

To account for parameter uncertainty, we can create distributions for key parameters:

```{r}
# Create parameter distribution for beta and gamma
param_distribution <- data.frame(
  beta = c(0.2, 0.3, 0.4, 0.5),
  gamma = c(0.1, 0.15, 0.2, 0.25),
  weights = c(0.25, 0.25, 0.25, 0.25)
)

# Set parameter distribution in the model
sir_model$set_param_dist(
  params_list = list(uncertainties = param_distribution),
  param_dist_weights = "weights",
  use_average = TRUE  # Use the average values instead of sampling
)

# View the parameter sets
head(sir_model$params_df)
```

## Saving and Loading Model Configuration

You can save your model configuration as JSON for later use:

```{r}
# Save model configuration as JSON
model_json <- sir_model$to_json()

# Create a new model and load configuration
new_sir_model <- SIRModel$new("restored_model")
new_sir_model$set_inputs_from_json(model_json)

# Verify model was loaded correctly
new_sir_model$inputs$beta
new_sir_model$inputs$gamma
```

## Conclusion

In this vignette, you've learned how to:

1. Create an epidemiological model using the `R6Sim` framework
2. Implement a classic SIR model with differential equations using `deSolve`
3. Run simulations and analyze results
4. Explore different parameter values
5. Set up parameter distributions for uncertainty analysis
6. Save and restore model configurations using JSON

In the next vignette, we'll demonstrate how to use the `R6Experiment` class to run systematic experiments with this SIR model, exploring different scenarios and parameter combinations.