---
title: "Running Experiments with R6Sim"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Running Experiments with R6Sim}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE
)
```

## Introduction

In the previous vignette, we created a basic SIR model using the `R6Sim` framework. Now, we'll explore how to use the `R6Experiment` class to run systematic experiments with this model, exploring different scenarios and parameter combinations.

## Setting Up Our SIR Model

First, let's recreate the SIR model from the previous vignette:

```{r setup}
library(R6Sim)
library(dplyr)
library(magrittr)
library(deSolve)

# Create an SIR epidemic model
SIRModel <- R6::R6Class(
  classname = "SIRModel",
  inherit = R6Sim,
  
  public = list(
    results = NULL,
    
    initialize = function(name) {
      super$initialize(name = name)
      
      # Set default parameter values
      self$set_input("N", 10000, "settings")$
        set_input("I0", 10, "settings")$
        set_input("R0", 0, "settings")$
        set_input("beta", 0.3, "natural_history")$
        set_input("gamma", 0.1, "natural_history")$
        set_input("t.max", 150, "settings")$
        set_input("dt", 1, "settings")
    },
    
    setup_run = function() {
      invisible(self)
    },
    
    sir_equations = function(time, state, parameters) {
      with(as.list(c(state, parameters)), {
        dS <- -beta * S * I / N
        dI <- beta * S * I / N - gamma * I
        dR <- gamma * I
        return(list(c(dS, dI, dR)))
      })
    },
    
    simulate = function() {
      initial_state <- c(
        S = self$inputs$N - self$inputs$I0 - self$inputs$R0,
        I = self$inputs$I0,
        R = self$inputs$R0
      )
      
      parameters <- c(
        N = self$inputs$N,
        beta = self$inputs$beta,
        gamma = self$inputs$gamma
      )
      
      times <- seq(0, self$inputs$t.max, by = self$inputs$dt)
      
      self$results <- as.data.frame(ode(
        y = initial_state,
        times = times,
        func = self$sir_equations,
        parms = parameters
      ))
      
      # Calculate epidemic metrics
      peak_infected <- max(self$results$I)
      peak_day <- self$results$time[which.max(self$results$I)]
      final_size <- self$results$R[length(self$results$R)]
      attack_rate <- final_size / self$inputs$N
      
      return(data.frame(
        peak_infected = peak_infected,
        peak_day = peak_day,
        final_size = final_size,
        attack_rate = attack_rate,
        R0 = self$inputs$beta / self$inputs$gamma
      ))
    }
  )
)

# Create a basic model instance
sir_model <- SIRModel$new("basic_sir")
```

## Creating an Experiment

Now, let's use the `R6Experiment` class to explore how different parameter combinations affect the epidemic outcomes:

```{r}
# Create a new experiment using our SIR model
experiment <- R6Experiment$new(sir_model)
```

## Defining Parameters for the Experiment

We'll define two key parameters to explore:

1. `beta` (transmission rate) - looking at different rates of disease spread
2. `gamma` (recovery rate) - exploring different disease durations

```{r}
# Define transmission rate parameter using a grid design
experiment$set_parameter(
  parameter_name = "beta",
  experimental_design = "grid",
  values = c(0.2, 0.3, 0.4, 0.5)
)

# Define recovery rate parameter using a grid design
experiment$set_parameter(
  parameter_name = "gamma",
  experimental_design = "grid",
  values = c(0.05, 0.1, 0.2)
)
```

## Setting Up the Experimental Design

Now we create the full experimental design by combining all parameter values:

```{r}
# Create the experimental design
experiment$set_design()

# Examine the resulting design
head(experiment$policy_design)
nrow(experiment$policy_design)
```

The experimental design now contains all combinations of our parameter values. Let's see the structure:

```{r}
# Show parameter combinations
experiment$policy_design %>%
  select(beta, gamma) %>%
  distinct()
```

## Running the Experiment

Now we'll run the experiment across all parameter combinations:

```{r}
# Run the experiment (in serial mode for simplicity)
results <- experiment$run(parallel = FALSE)

# Examine the first few results
head(results)
```

## Analyzing the Results

Now that we have our results, we can analyze how different parameters affect the epidemic outcomes:

```{r, fig.width=7, fig.height=5}
# Summary statistics by beta and gamma
summary_results <- results %>%
  group_by(beta, gamma) %>%
  summarize(
    mean_peak_infected = mean(peak_infected),
    mean_peak_day = mean(peak_day),
    mean_attack_rate = mean(attack_rate),
    mean_R0 = mean(R0),
    .groups = "drop"
  )

summary_results
```

Let's visualize how the transmission rate (beta) and recovery rate (gamma) affect the peak number of infected individuals:

```{r, fig.width=7, fig.height=5}
# Visualization if ggplot2 is available
if (requireNamespace("ggplot2", quietly = TRUE)) {
  library(ggplot2)
  
  ggplot(summary_results, aes(x = beta, y = gamma, fill = mean_peak_infected)) +
    geom_tile() +
    scale_fill_viridis_c() +
    theme_minimal() +
    labs(title = "Peak Infected Population by Transmission and Recovery Rates",
         x = "Transmission Rate (β)",
         y = "Recovery Rate (γ)",
         fill = "Peak Infected")
}
```

## Adding Stochasticity with Replications

To account for stochasticity in our model, we can use replications:

```{r}
# Create a new experiment with a stochastic SIR model
StochasticSIRModel <- R6::R6Class(
  classname = "StochasticSIRModel",
  inherit = SIRModel,
  
  public = list(
    # Override simulate to add stochasticity
    simulate = function() {
      # Call parent's simulate method first
      result <- super$simulate()
      
      # Add some stochasticity to the results (simplified example)
      noise_factor <- rnorm(1, mean = 1, sd = 0.05)
      result$peak_infected <- result$peak_infected * noise_factor
      result$final_size <- result$final_size * noise_factor
      
      return(result)
    }
  )
)

# Create stochastic model
stochastic_model <- StochasticSIRModel$new("stochastic_sir")

# Create experiment with stochastic model
stochastic_experiment <- R6Experiment$new(stochastic_model)

# Set parameters (using fewer combinations for simplicity)
stochastic_experiment$set_parameter(
  parameter_name = "beta",
  experimental_design = "grid",
  values = c(0.3, 0.4)
)

stochastic_experiment$set_parameter(
  parameter_name = "gamma",
  experimental_design = "grid",
  values = c(0.1, 0.2)
)

# Set up design with 5 replications for each parameter combination
stochastic_experiment$set_design(n_reps = 5, set_seed = TRUE)

# Examine the design - note the rep.id and seed columns
head(stochastic_experiment$policy_design)
```

Now we can run the stochastic experiment:

```{r}
# Run the stochastic experiment
stochastic_results <- stochastic_experiment$run(parallel = FALSE)

# Examine results with replications
stochastic_results %>%
  group_by(beta, gamma) %>%
  summarize(
    mean_peak_infected = mean(peak_infected),
    sd_peak_infected = sd(peak_infected),
    cv_peak_infected = sd_peak_infected / mean_peak_infected,
    .groups = "drop"
  )
```

## Comparing Multiple Model Variants

One powerful feature of `R6Experiment` is the ability to run experiments with multiple model variants. Let's create a modified SIR model with vaccination:

```{r}
# Create a simpler comparison example with two standard models
model_A <- SIRModel$new("model_A")
model_B <- SIRModel$new("model_B")

# Set different default parameters for model B
model_B$set_input("beta", 0.4, "natural_history")
model_B$set_input("gamma", 0.15, "natural_history")

# Create an experiment with both models
comparison_experiment <- R6Experiment$new(model_A, model_B)

# Set experimental parameters
comparison_experiment$set_parameter(
  parameter_name = "beta", 
  experimental_design = "grid",
  values = c(0.3, 0.5)
)

# Generate the experimental design
comparison_experiment$set_design()

# Run the experiment
comparison_results <- comparison_experiment$run(parallel = FALSE)

# Analyze results by model and beta
comparison_results %>%
  group_by(model.id, beta) %>%
  summarize(
    mean_peak_infected = mean(peak_infected),
    mean_attack_rate = mean(attack_rate),
    .groups = "drop"
  )
```

## Selecting Specific Parameter or Design Points

You can also run experiments with specific parameter sets or design points:

```{r}
# Let's run the experiment again, but only for the first 2 parameter combinations
selected_results <- experiment$run(
  parallel = FALSE
)[1:2, ]

# View the results
selected_results
```

## Conclusion

In this vignette, we've demonstrated how to use the `R6Experiment` class to:

1. Set up structured experiments with multiple parameters
2. Run simulations across all parameter combinations
3. Add stochasticity with replications
4. Compare multiple model variants
5. Analyze and visualize experiment results
6. Select specific design points for simulation

The `R6Experiment` framework provides a powerful way to systematically explore model behavior across different parameter settings and scenarios, enabling comprehensive sensitivity analysis and policy evaluation.