---
title: "Introduction to R6Sim"
output: rmarkdown::html_vignette
author: "Pedro Nascimento de Lima"
vignette: >
  %\VignetteIndexEntry{Introduction to R6Sim}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This package contains two `R6` classes that can be inherited in your analysis: `R6Sim` and `R6Experiment`. Here I provide a brief introduction of how they work. A more complete example of their use can be seen in repositories that build models and analyses based on `R6Sim`, such as the [value of environmental surveillance](github.com/randcorporation/value-of-surveillance) project.

### The `R6Sim` model class

The `R6Sim` class is meant to represent an R6 simulation model. An `R6Sim` model should be self-contained, and once instantiated, should immediately know how to simulate itself with the `model$simulate()` function. Here is an example `R6Sim` model:

```{r}
library(dplyr)
library(R6Sim)

Mymodel <- R6::R6Class(
  classname = "Mymodel",
  inherit = R6Sim,
  public = list(

    sim_res = NULL,

    # Custom Initialize function
    initialize = function(name) {
      super$initialize(name = name)
      self$set_input("pop.size", 100)$
           set_input("risk.mean", 0.01)$
           set_input("risk.sd", 0.001)$
           set_input(name = "trials",value = 10)
      
      # Can also add a parameter distribution 
      self$set_param_dist(params_list = list(param_dist_a = data.frame(sample_param = 1, weights = 1)), use_average = T, param_dist_weights = "weights")

    },

    # Sample Simulate function
    simulate = function() {

      # Create a sample population with some health events:
      self$sim_res = data.frame(p.id = 1:self$inputs$pop.size,
                           risk = rnorm(n = self$inputs$pop.size, mean = self$inputs$risk.mean, sd = self$inputs$risk.sd)) %>%
        mutate(probability.event = 1-exp(-risk)) %>%
        mutate(n.events = rbinom(n = 1:self$inputs$pop.size, size = self$inputs$trials, prob = probability.event))
      
      # Let's return the number of events as the key model input
      return(data.frame(events = sum(self$sim_res$n.events)))
    }
  )
)

model <- Mymodel$new("test model")

model$simulate()
```

Once your model is created, you can set inputs and run the model again:

```{r}
# you can chain methods
model$set_input("risk.mean", 0.05)$
  simulate()
```

The `model$inputs` list contains all model inputs, and can be inspected. 
```{r}
model$inputs
```
Instead of passing these inputs around as parameters in your functions, you can always access the model inputs whithin the model with the `self$inputs$input_name`. 

### The `R6Experiment` class

An `R6Experiment` can contain multiple `R6Sim` models. That class can be used to define experiments and run them in parallel. Your experiment can have multiple models.

```{r}

# Now, the other model will be slightly different
my_other_model <- model$clone(deep = T)
my_other_model$
  set_input("risk.sd", 0.002)

my_other_model$name <- "high dispersion"

# Note that my_other_model could be an arbitraryly different class, as long as it also inherits R6Sim

# here, we could change some structural parameter:

experiment = R6Experiment$new(model, my_other_model)

experiment$
  set_parameter(parameter_name = "risk.mean",experimental_design = "grid",values = seq.default(from = 0.01, to = 0.05, by = 0.01))$
  set_design()

results <- experiment$run(parallel = F, model_from_cluster_eval = F)

# model.id =1 is the first model passed to experiment, and so on:
results %>%
  select(model.id, risk.mean, events)

```

Once you experiment runs, now you can iterate and add new inputs, or new parameters to your experiment, or add an arbitrary number of models to your experiment. To see a real use case with parallel runs, see the [value-of-surveillance](https:://github.com/randcorporation/value-of-surveillance) project.
